Studying the Performance of the Jellyfish Search Optimiser for the Application of Projection Pursuit

### 2024-09-04 jso.qmd
working on JSO in PPGT

>>>>

(not done yet)

The JSO can be used by specifying the argument `search_f = search_jellyfish` in the `guided_tour()` function in the `tourr` package. In the example below, the `lda_pp` index is used to cluster the penguins data set. When binding animation `animate_xy()` to an object (`res<-`), it will save the visited bases of each jellyfish as a data object. The bases from each jellyfish can be extracted adn supplied as a planned tour to visualise its path. Figure @fig-jso-penguins shows the tour path of one particular jellyfish in the penguins data set. The jellyfish successfully finds the structure of the data set, separating the three species of penguins.

```{r}
#| echo: true
#| eval: false
# run the projection pursuit with JSO 
res <- animate_xy(penguins_df, 
                  guided_tour(lda_pp(cl = penguins_cl$species),
                              search_f = search_jellyfish, 
                              n_jellies = 20, max.tries = 50))
# extract the bases from the first jellyfish
bases <- res |> filter(loop == 1) |> pull(basis) |> check_dup(0.1)
# visualise its path with a planned tour
animate_xy(data = penguins_df, tour_path = planned_tour(bases), 
           col = penguins_cl$species)
```

```{r}
#| echo: false
#| eval: false
library(dplyr)
library(palmerpenguins)
library(ggplot2)
library(tourr)
check_dup <- function(bases, min_dist) {
  res <- list()
  res[[1]] <- bases[[1]]
  i <- 1; j <- 2
  while(j <= length(bases)){
    if (proj_dist(bases[[i]], bases[[j]]) >= min_dist) {
      res <- c(res, list(bases[[j]]))
      i <- i + 1; j <- j + 1
    } else{
      j <- j + 1
    }
  }
  return(res)
}

penguins_cl <- palmerpenguins::penguins |> filter(!is.na(bill_length_mm)) |>
  rename(bl = bill_length_mm, bd = bill_depth_mm,
         fl = flipper_length_mm, bm = body_mass_g)
penguins_df <- penguins_cl |> select(bl:bm) |> rescale()
res <- animate_xy(penguins_df, guided_tour(lda_pp(cl = penguins_cl$species), search_f = search_jellyfish, n_jellies = 20, max.tries = 50))
bases <- res |> filter(loop == 1) |> pull(basis) |> check_dup(0.1)
animate_xy(data = penguins_df, tour_path = planned_tour(bases), col = penguins_cl$species)
render_gif(penguins_df, planned_tour(bases), display_xy(col = penguins_cl$species), 
           here::here("figures/tour/penguins.gif"))
render( 
  penguins_df,
  tour_path = planned_tour(bases),
  dev = "png",
  display = display_xy(col = penguins_cl$species),
  frames = 100,
  here::here("figures","tour", "penguins%03d.png")
)
```

```{r}
#| label: fig-jso-penguins
#| fig-cap: "Visualising the tour path of one particular jellyfish in the penguins data set. The jellyfish is optimised by the JSO with 20 jellyfishes and a maximum of 50 tries. The tour path is planned off-line and then followed in the guided tour. The jellyfish successfully finds the structure of the data set, separating the three species of penguins."
frames <- c("001", "010", "019", "028", "036", "048")
ani <- paste0(here::here("figures/"), "tour/", "penguins", frames, ".png")
rl <- lapply(ani, png::readPNG)
gl <-  lapply(rl, grid::rasterGrob)
wrap_plots(gl)
```

### 2024-09-04 jso.tex
working on JSO in PPGT

\begin{tcolorbox}[enhanced jigsaw, coltitle=black, rightrule=.15mm, colframe=quarto-callout-note-color-frame, left=2mm, colbacktitle=quarto-callout-note-color!10!white, leftrule=.75mm, breakable, opacitybacktitle=0.6, arc=.35mm, title={Algorithm: Jellyfish Optimizer Pseudo Code}, colback=white, opacityback=0, bottomtitle=1mm, titlerule=0mm, toptitle=1mm, bottomrule=.15mm, toprule=.15mm]

@@

>>>>

\begin{tcolorbox}[enhanced jigsaw, leftrule=.75mm, title={Algorithm: Jellyfish Optimizer Pseudo Code}, colback=white, bottomrule=.15mm, arc=.35mm, breakable, colbacktitle=quarto-callout-note-color!10!white, rightrule=.15mm, bottomtitle=1mm, toptitle=1mm, colframe=quarto-callout-note-color-frame, toprule=.15mm, coltitle=black, opacitybacktitle=0.6, left=2mm, opacityback=0, titlerule=0mm]

@@

(not done yet)

The JSO can be used by specifying the argument
\texttt{search\_f\ =\ search\_jellyfish} in the \texttt{guided\_tour()}
function in the \texttt{tourr} package. In the example below, the
\texttt{lda\_pp} index is used to cluster the penguins data set. When
binding animation \texttt{animate\_xy()} to an object
(\texttt{res\textless{}-}), it will save the visited bases of each
jellyfish as a data object. The bases from each jellyfish can be
extracted adn supplied as a planned tour to visualise its path. Figure
Figure~\ref{fig-jso-penguins} shows the tour path of one particular
jellyfish in the penguins data set. The jellyfish successfully finds the
structure of the data set, separating the three species of penguins.

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\# run the projection pursuit with JSO }
\NormalTok{res }\OtherTok{\textless{}{-}} \FunctionTok{animate\_xy}\NormalTok{(penguins\_df, }
                  \FunctionTok{guided\_tour}\NormalTok{(}\FunctionTok{lda\_pp}\NormalTok{(}\AttributeTok{cl =}\NormalTok{ penguins\_cl}\SpecialCharTok{$}\NormalTok{species),}
                              \AttributeTok{search\_f =}\NormalTok{ search\_jellyfish, }
                              \AttributeTok{n\_jellies =} \DecValTok{20}\NormalTok{, }\AttributeTok{max.tries =} \DecValTok{50}\NormalTok{))}
\CommentTok{\# extract the bases from the first jellyfish}
\NormalTok{bases }\OtherTok{\textless{}{-}}\NormalTok{ res }\SpecialCharTok{|\textgreater{}} \FunctionTok{filter}\NormalTok{(loop }\SpecialCharTok{==} \DecValTok{1}\NormalTok{) }\SpecialCharTok{|\textgreater{}} \FunctionTok{pull}\NormalTok{(basis) }\SpecialCharTok{|\textgreater{}} \FunctionTok{check\_dup}\NormalTok{(}\FloatTok{0.1}\NormalTok{)}
\CommentTok{\# visualise its path with a planned tour}
\FunctionTok{animate\_xy}\NormalTok{(}\AttributeTok{data =}\NormalTok{ penguins\_df, }\AttributeTok{tour\_path =} \FunctionTok{planned\_tour}\NormalTok{(bases), }
           \AttributeTok{col =}\NormalTok{ penguins\_cl}\SpecialCharTok{$}\NormalTok{species)}
\end{Highlighting}
\end{Shaded}

\includegraphics{jso_files/figure-pdf/fig-jso-penguins-1.pdf}

\caption{\label{fig-jso-penguins}Visualising the tour path of one
particular jellyfish in the penguins data set. The jellyfish is
optimised by the JSO with 20 jellyfishes and a maximum of 50 tries. The
tour path is planned off-line and then followed in the guided tour. The
jellyfish successfully finds the structure of the data set, separating
the three species of penguins.}
